Online {\IterComp}

### 2017-09-13 motivation.tex
minor
motivation first part

%As a motivation example, consider...

%Here we give an example to motivate our work. E.g. assuming we have a zero-cost profiling based work metric,
%how can we use it to perform iterative compilation across runs? Then we talk about while it is impossible to
%build a zero-overhead metric, we can develop one with low overhead.

This section provides a comparison between the offline and online approaches for
{\itercomp}.
The goal is to motivate the need for an online {\itercomp}, highlighting
weaknesses of using offline {\itercomp}.

%In the offline approach, a fixed set of inputs is selected for which the best
%optimization sequence is selected during the development time (pre-shipping) of
%a program.
%Although this fixed set of inputs is expected to be representative of the user's
%workload,
%in most cases, it might be infeasible to identify and collect representative
%workloads for every group of users.
%Furthermore, a user can dynamically change its workload pattern, rendering
%a fixed optimization sequence innapropriate for his performance requirements.

%In the online approach, the program is shipped with an initial optimization
%sequence and different optimization sequences are evaluated as the end-user
%executes the program.
%This approach frees developers from the need to collect representative
%workloads for every group of users so that the program can be optimized
%accordingly.
%More importantly, even in scenarios where the user's workload can dynamically
%change its pattern, the online {\itercomp} would still be able to optimize the
%program accordingly.

In the offline approach, a small fixed set of inputs is selected for which all
optimization sequences are evaluated.
In the online approach, each optimization sequence is evaluated based on a,
dynamically selected, distinct set of inputs.
In both cases, the optimization sequences are ranked based on their average
speedup, and the best optimization sequence is selected.

In order to assess the quality of the optimization sequence selected by each
approach, we evaluate both optimization sequences against all available inputs
in our dataset.
Figure~\ref{fig:motivation-online} shows this comparison between both the offline
and online approaches for {\itercomp}.
It shows the performance gains of both optimization sequences compared with the
standard compiler's optimization (i.e.,\texttt{-O3}).
The online {\itercomp} achieves an average speedup which is about 13\% better
than the solution selected by the offline approach.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
    \caption{Comparison between the offline and online {\itercomp}.
             The vertical lines represent the average speedups over
             all the 1,000 inputs.}
    \label{fig:motivation-online}
\end{figure}

It is worth mentioning that our goal is to improve the average performance
across the user's workload, and not only for a specific input.
This means that the online {\itercomp} is better informed throughout the search,
as it compares optimization sequences across different sets inputs.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-metric.pdf}
    \caption{Comparison among different performance metrics for performing
             online {\itercomp}.
             The vertical lines represent the best score for each metric.}
    \label{fig:motivation-metric}
\end{figure}

>>>>

Offline \itercomp can be very sensitive on having representative inputs for the program under optimization. If the
developers cannot predict what will be the typical usage of the application or that typical usage changes over time,
then \itercomp might fail making the application faster for the user.

To show this, we randomly generated multiple
input subsets out of the full set of XXX inputs. Working with each subset, \itercomp found a ``best'' optimization
sequence which maximized the speedup for the given input subset. We evaluated the true quality of each such sequence
by measuring the average speedup against \texttt{-O3} for all the XXX inputs. The green bars in
Figure~\ref{fig:motivation-online} show the histogram of those speedups. Depending on the selection of inputs,
performance can degrade by 18\% compared to \texttt{-O3}, improve by 46\%, or anything in between. Overall, almost half
of the optimization sequences discovered fail to improve performance. This means that without intelligent selection of
inputs, offline \itercomp might be a waste of time.

The blue bars in Figure~\ref{fig:motivation-online} show a different approach where during \itercomp the speedup of
each optimization sequence is again measured on a small input subset but this subset changes for every optimization
sequence. This is similar to how online \itercomp would have to work, using each input only once. This might introduce
errors, since the speedups we measure for each optimization sequence are not strictly comparable, they are measured on
different workloads. But in the end the histogram shows that being able to test more inputs allows us to make better
informed decisions. This approach never degrades performance compared to \textt{-O3}, performance varies less, and on
average is 13\% better than the offline approach.

% %As a motivation example, consider...
% 
% %Here we give an example to motivate our work. E.g. assuming we have a zero-cost profiling based work metric,
% %how can we use it to perform iterative compilation across runs? Then we talk about while it is impossible to
% %build a zero-overhead metric, we can develop one with low overhead.
% 
% This section provides a comparison between the offline and online approaches for
% {\itercomp}.
% The goal is to motivate the need for an online {\itercomp}, highlighting
% weaknesses of using offline {\itercomp}.
% 
% %In the offline approach, a fixed set of inputs is selected for which the best
% %optimization sequence is selected during the development time (pre-shipping) of
% %a program.
% %Although this fixed set of inputs is expected to be representative of the user's
% %workload,
% %in most cases, it might be infeasible to identify and collect representative
% %workloads for every group of users.
% %Furthermore, a user can dynamically change its workload pattern, rendering
% %a fixed optimization sequence innapropriate for his performance requirements.
% 
% %In the online approach, the program is shipped with an initial optimization
% %sequence and different optimization sequences are evaluated as the end-user
% %executes the program.
% %This approach frees developers from the need to collect representative
% %workloads for every group of users so that the program can be optimized
% %accordingly.
% %More importantly, even in scenarios where the user's workload can dynamically
% %change its pattern, the online {\itercomp} would still be able to optimize the
% %program accordingly.
% 
% In the offline approach, a small fixed set of inputs is selected for which all
% optimization sequences are evaluated.
% In the online approach, each optimization sequence is evaluated based on a,
% dynamically selected, distinct set of inputs.
% In both cases, the optimization sequences are ranked based on their average
% speedup, and the best optimization sequence is selected.
% 
% In order to assess the quality of the optimization sequence selected by each
% approach, we evaluate both optimization sequences against all available inputs
% in our dataset.
% Figure~\ref{fig:motivation-online} shows this comparison between both the offline
% and online approaches for {\itercomp}.
% It shows the performance gains of both optimization sequences compared with the
% standard compiler's optimization (i.e.,\texttt{-O3}).
% The online {\itercomp} achieves an average speedup which is about 13\% better
% than the solution selected by the offline approach.

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
    \caption{Comparison between the offline and online {\itercomp}.
             The vertical lines represent the average speedups over
             all the 1,000 inputs.}
    \label{fig:motivation-online}
\end{figure}

% It is worth mentioning that our goal is to improve the average performance
% across the user's workload, and not only for a specific input.
% This means that the online {\itercomp} is better informed throughout the search,
% as it compares optimization sequences across different sets inputs.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-metric.pdf}
    \caption{Comparison among different performance metrics for performing
             online {\itercomp}.
             The vertical lines represent the best score for each metric.}
    \label{fig:motivation-metric}
\end{figure}
